\begin{problem}
  Formally specify a DFA for the set of all strings in $\Sigma = \set{0, 1}^*$
  such that each string is of length at least $2000$ and every block of $2000$
  consecutive symbols has at least ten $0$s.
  Explain the design in one or two sentences.
\end{problem}

\begin{Answer}
  \step
  \textbf{DFA Specification}

  \step
  To simplify the expressions, we define the following functions:
  \begin{align*}
    \shift{s} &= \begin{cases}
      \eps &\text{ if } s = \eps\\
      x &\text{ otherwise, where } s \text{ is parsed as } \crim{\set{0, 1}^1x} \\
    \end{cases} \\
    \beta(s) &= n \text{ where $n$ is the integer obtained by interpreting $s$ as a binary number.} \\
    \#(s) &= n \text{ where $n$ is the length of $s$.} \\
    \#_a(s) &= n \text{ where $n$ is the number of occurrences of $a$ in $s$.}
  \end{align*}
  \step
  We specify the DFA as:
  \begin{align*}
    M &= \left( Q,\ \Sigma,\ \delta,\ q_0,\ F \right) \text{ where: }\\
    Q &= \set{ q_{(i,\ s)} : 0 \leq i \le 2000,\; s \in \set{0, 1}^* \text{ and } \#(s) = i} \\
    \Sigma &= \set{0, 1} \\
    \delta(q_{(i,\ s)}, x) &= \begin{cases}
      q_{(i+1,\ sx)} &\text{ if } i < 2000 \quad \quad \quad \quad \quad \quad \quad \quad \zaff{\text{// Still too short.}}\\
      q_{(i, \shift{s}x)} &\text{ if $i = 2000$ and } \#_0(s) \ge 10 \quad \zaff{\text{ // Valid length and condition.}} \\
      q_{(i,\ s)} &\text{ otherwise.} \quad \quad \quad \quad \quad \quad \quad \quad \zaff{\text{// Invalid condition, trap in the state.}}
    \end{cases} \\
    q_0 &= q_{(0,\ \eps)} \\
    F &= \set{ q_{(2000,\ s)} : \#_0(s) \ge 10}
  \end{align*}

  \newpage
  \step
  \textbf{Main Idea}:

  \step
  We track states as a tuple of the length of the string processed so far (say, $n$)
  and the most recent $n$ characters processed from the string.
  While $n$ is less than $2000$, each transition appends the current character
  to the right of the string we are tracking and increments $n$.
  Once $n$ reaches $2000$, we no longer
  need to keep incrementing it since the length requirement is satisfied.
  We also start watching for and trapping at invalid states.
  While transitioning in valid states, we drop the left-most character
  of the string we are tracking and append the new character to the right.

  \step
  Finally, we accept if we have processed at least $2000$ characters
  (i.e. we are at a state $q_{(\crim{2000},\ s)})$
  and the latest $2000$ characters have at least ten $0$
  (i.e. $\#_0(s) \ge 10$).
\end{Answer}
