\begin{problem}
  For each of the following languages, decide whether it is regular
  or not. If it is regular, give a finite automaton or a regular expression for it.
  If is is not regular, use any combination of
  the pumping lemma, closure properties, or results \emph{proved} in class.
  
  \step
  The pumping lemma stipulates that:
  \begin{lemma}
    Pumping lemma for regular languages.

    \step
    \emph{Let $L \subseteq \Sigma^*$ be a regular language over an alphabet $\Sigma$.
    Then, there exists a positive integer $p$, called the
    pumping length, such that for all $s \in L, \abs{s} > p$,
    there exists $u, v, w \in \Sigma^*$ such that:
    \begin{enumroman}
      \item $s = uvw$
      \item $\abs{uv} \le p$
      \item $\abs{v} > 0$
      \item $uv^iw \in L$ for all $i \geq 0$
    \end{enumroman}
    }
  \end{lemma}

  \begin{enumalph}
    \item $L_1 = \set{0^m 1^n 0^{m + n} : m, n \geq 0}$.
      \begin{Answer}
        The language is not regular.

        \step
        Assuming $L_1$ was regular and $p$ was the pumping length for $L_1$,
        take $s = 0^p1^k0^{p+k}$ for some $k \geq 0$.
        Better yet, set $k = 0$, so that $s = 0^p0^{p}$.
        Following the pumping lemma, take $u = 0^a, v = 0^b, w = 0^{p-a-b}0^{p}$.
        The pumping lemma stipulates that:
        \begin{enumroman}
          \item $\abs{uv} \le p$.
          \item $\abs{v} = b > 0$.
        \end{enumroman}
        
        \step
        If we pump down the string, we see then $uw \in L_1$,
        so $0^a0^{p-a-b}0^{p} = 0^{p-b}0^{p} \in L_1$.
        Therefore, $p - b = p$ (by definition of $L_1$), implying that $b = 0$.
        However, the pumping stipulated that $b > 0$, so we have a contradiction,
        meaning $L_1$ may not be regular.
      \end{Answer}
    \item $L_2 = \set{xwx^R : x, w \in \set{0, 1}^*, \abs{x} > 0 \text{ and } \abs{y} > 0}$.
      \begin{Answer}
        The language is not regular.

        \step
        Assuming $L_2$ was regular and $p$ was the pumping length for $L_2$,
        Take $s = 0^p10^p$.
        Following the pumping lemma, take $u = 0^a, v = 0^b, w = 0^{p-a-b}10^p$.
        The pumping lemma stipulates that:
        \begin{enumroman}
          \item $\abs{uv} \le p$.
          \item $\abs{v} = b > 0$.
        \end{enumroman}
        If we pump down the string, we see then $uw \in L_2$,
        so $0^a0^{p-a-b}10^p = 0^{p-b}10^p \in L_2$.
        This implies that $p - b = p$, so $b = 0$,
        contradicting the pumping lemma.
        Therefore, $L_2$ may not be regular.
      \end{Answer}
    \item $L_3 = \set{0^m 1^n : m \text{ divides } n}$.
      \begin{Answer}
        The language is not regular.

        \step
        Assuming $L_3$ was regular and $p$ was the pumping length for $L_3$,
        Take $s = 0^p1^p$ so that $s \in L_3$.
        Following the pumping lemma, take $u = 0^a, v = 0^b, w = 0^{p-a-b}1^p$.
        The pumping lemma stipulates that:
        \begin{enumroman}
          \item $\abs{uv} \le p$.
          \item $\abs{v} = b > 0$.
        \end{enumroman}
        If we pump up the string, we see then $uv^2w \in L_2$,
        so $0^a0^{2b}0^{p-a-b}1^p = 0^{p+b}1^p \in L_2$.
        This implies that $p + b$ divides $p$, so $b = 0$,
        contradicting the pumping lemma.
        Therefore, $L_3$ may not be regular.
      \end{Answer}
    \item $L_4 = \set{0^n 1^p : n \le 4 \text{ or $p$ is prime (or both)}}$.
      \begin{Answer}
        First, note that non-regular languages are closed under reversal,
        i.e. if $L$ is not regular, then $L^R$ is also not regular.
        The intuition is as follows:
        Given a regular language $L$ with a DFA $M = (Q, \Sigma, \delta, q_0, F)$ that recognizes $L$,
        we can easily construct an NFA for $L_R$ as follows:
        \begin{enumroman}
          \item Reverse all the transitions in $M$.
          \item Add a new start state $s \notin Q$ with $\eps$-transitions
          from $s$ to every state $f \in F$.
          \item Change all the accepting states to non-accepting states.
          \item Change the old start state to not be a start state, but make it an accepting state.
        \end{enumroman}

        \step
        \begin{claim}
          $L_4^R$ is not regular, therefore $L_4$ is not regular.

          \begin{proof}
            Note that $L_4^R = \set{1^m0^n : \text{$m$ is prime and $n \ge 4$}}$.

            \step
            Suppose $p$ is the pumping length for $L_4^R$, pick $s = 1^p0^5$.
            By the pumping lemma, take $u = 1^a, v = 1^b, w = 1^{p-a-b}0^5$.
            The pumping lemma stipulates that $uv^kw \in L_4^R$ for all $k \geq 0$.
            However, note that:
            \begin{align*}
              uv^kw &= 1^a1^{kb}1^{p-a-b}0^5 \in L_4^R\\
              &\implies 1^{p+(k-1)b}0^5 L_4^R\\
              &\implies p+(k-1)b \text{ is prime.}
            \end{align*}
            Set $k = p+1$, then $p+(k-1) = p + pb = p(1+b)$.
            However, $b > 0$ by the pumping lemma, so $p(1+b)$ cannot be prime.
            This contradiction implies that $L_4^R$ is not regular.
            Therefore, $L_4$ cannot be regular.
          \end{proof}
        \end{claim}
      \end{Answer}
    \item The infinite union $\bigcup_{n \geq 1}^\infty A_i$, where each $A_i$
      is a regular language.
      \emph{
        The question should be interpreted as asking
        whether the union is \textbf{guaranteed} to be regular
        no matter how the sets $A_i$ are chosen.
      }
      \begin{Answer}
        No, the union is not guaranteed to be regular.
        For a counter-example, set $A_i = 0^i1^i$ for all $i \geq 1$.
        Then, $\bigcup_{n \geq 1}^\infty A_i = \set{0^n1^n : n \ge 1}$,
        which is not a regular language.
      \end{Answer}
    \item The infinite intersection $\bigcap_{n \geq 1}^\infty A_i$, where each $A_i$
      is a regular language.
      \emph{
        The question should be interpreted as asking
        whether the intersection is \textbf{guaranteed} to be regular
        no matter how the sets $A_i$ are chosen.
      }
      \begin{Answer}
        No, the language is not guaranteed to be regular.
        For a counter-example:
        Let $P' = \set{0, 1, 4, 6, 8, 9, 10, \ldots}$ be the set of all composite numbers.
        Let $P_i$ refer to the $i$'t element in the sequence.
        
        \step
        Define $\Sigma = \set{1}$ and $A_i = 1^* - \set{1^{P_i}}$ for all $i \geq 1$.
        Then, define the language $L$ as follows:
        \[ L = \bigcap_{n \geq 1}^\infty A_i = \set{1^p : \text{ $p$ is prime. }} \]

        \step
        As we proved in class, the language $L$ is not regular.
      \end{Answer}
  \end{enumalph}

\end{problem}

