\begin{problem}
  Now, apply the notion of left equivalence to the theory
  of regular languages.

  \begin{enumalph}
    \item For a language $A$ over alphabet $\Sigma$ and a string $x \in \Sigma$,
      let $[x]_A$ denote the equivalence class (of $A$) to which $x$ belongs.
      For instance, consider the set $X_1 = (01)^*, X_2 = (01)^*0$,
      and $X_3 = \set{0, 1}^* - (X_1 \cup X_2)$, then:
      \begin{enumroman}
        \item $[010]_{B}$ denotes the set $X_2$,
        \item $[01010]_{B}$ also denotes the same set $X_2$, and
        \item $[\eps]_B$ and $[0101]_B$ both denote the set $X_1$.
      \end{enumroman}

      \step
      There are rarely one unique way way to write an equivalence class
      $[x]_A$ --- there are usually multiple choices for $x$.
      These choices are called \emph{representatives}
      of the equivalence class.

      \step
      Prove that for any $x \in \Sigma^*$ and $a \in \Sigma$,
      the class $[xa]_A$ is completely determined by the class $[x]_A$
      and the alphabet symbol $a$ ---
      that is, prove that the particular $x$ we pick as a representative
      for the class $[x]_A$ is inconsequential.
      \begin{Answer}
        Recall the definition of equivalence classes:
        two strings $x, y$ are in the same equivalence class iff
        $x \equiv_A y$, i.e. $xw \in A \iff yw \in A$ for all $w \in \Sigma^*$.
        Consequently, we have that if $xw \in A$ for any string $x$ in $A$,
        then $yw \in A$ for every other element $y$ in the equivalence class.
        Therefore, the choice of a member of the equivalence class as a
        representative does not uniquely determine the set of transitions
        since every other member of the equivalence class would make the
        same set of accepting and rejecting transitions.
      \end{Answer}

    \item Suppose a language $A$ over alphabet $\Sigma$ has finitely many
      equivalence classes; $[x_1]_A, [x_2]_A, \ldots, [x_n]_A$ for some
      $n \ge 1$. Prove that $A$ is regular.
      \begin{Answer}
        Let $A$ be a language with finitely many equivalence classes.
        As shown in part (a) above, for any equivalence class
        $[x]_A$, the equivalence class $[xa]_A$ is not dependent on the
        choice of $x$ as a representative.
        Therefore, we can capture each unique equivalence class
        $[x]_A$ with a single state in an automaton and,
        since $A$ has finitely many equivalence classes,
        we can construct an automaton with finite states that accepts $A$.
      \end{Answer}
  \end{enumalph}
\end{problem}
