
\def \diverse { \textsc{Diverse} }
\def \indset { \textsc{IndSet} }
\begin{problem}
  You are in charge of analyzing a large social media network modeled as an
  undirected graph $G = (V, E)$: each vertex is a user and each edge
  represents a mutual ``friendship'' relation between two users.
  A \emph{diverse sample} is a set of users such that no two o them are friends
  and, furthermore, no two of them have a friend in common. Formally, a set
  $S \subseteq V$ is a diverse sample if
  \[ 
    \forall x, y \in S : \set{x, y} \notin E \land \nexists w \in V
    ( \set{w, x} \in E \land \set{w, y} \in E).
  \]
  As part of your analysis process, you have been asked to determine the largest
  diverse sample in the social network. After failing to find an efficient algorithm
  for this, you begin to suspect that this might be impossible.
  Prove, assuming that $\sfP \neq \sfNP$, there is indeed no polynomial-time
  algorithm for this optimization problem.
\end{problem}
\begin{Answer}
  First, let us define the language
  \[
    \diverse :=
    \set{ \vector{G, k} : \text{$G$ is a graph and $G$ has a diverse
      sample of size $k$}}.
  \]
  If $\diverse$ were decidable, then the problem of finding the largest possible
  diverse sample could be solved in polynomial time as follows:

  \step
  $T_0$ = ``On input $\vector{G}$, where $G$ is a graph:
  \begin{enumarabic}
    \item $S \gets \emptyset$.
    \item for $i \gets 1 \ldots \abs{V}$
      \begin{enumarabic*}
        \item if $\vector{G, k} \in \diverse$:
        \item $S \gets$ diverse sample of size $i$.
      \end{enumarabic*}
    \item \Return $S$.
  \end{enumarabic}

  \step
  Thus, we show that $\diverse$ is \sfNP-complete.

  \diverse can be solved in \sfNP by guessing a subset $S \subset 2^V$ and
  checking if the subset is diverse.

  We proved in class that the set
  \[ \indset := \set{\vector{G, k} : G \text{ has an independent set of size } k} \]
  is \sfNP-complete. We can reduce \indset to \diverse as follows:
  
  \newpage
  $T_1$ = ``On input $\vector{G, k}$, where $G$ is a graph and $k$ is an integer:
  \begin{enumarabic}
    \item Define the set
      $E' = E \setminus \set{ \set{u, w}, \set{v, w} : u, v, w \in V, \set{u, v} \not \in E,
      \set{\set{u, w}, \set{v, w}} \subseteq E}$.
    \item Define the graph $G' = (V, E')$.
    \item Output $\vector{G', k}$.''
  \end{enumarabic}

  \step
  \begin{align*}
    \vector{G, k} \in \indset
      &\Iff \exists S \subseteq V, \abs{S} = k \text{ and } \set{u, v} \not \in E
        \text{ for all } u, v \in S \\
      &\Iff \set{u, v} \notin E' \text{ and } \nexists w \in V
      ( \set{w, x} \in E' \land \set{w, y} \in E') \\
      &\Iff \vector{G', k} \in \diverse
  \end{align*}

  Since $\indset$ is \sfNP-complete, \diverse must also be \sfNP-complete.
  However, if $\diverse \in \sfNP$ and $\sfP \neq \sfNP$, then
  $\diverse \not \in \sfP$ and there cannot exist a polynomial time algorithm
  for finding the largest diverse sample.
\end{Answer}


