\begin{problem}
  Show that one of the languages $\EqCFG$ and $\EqCFGcomp$
  is recognizable, whereas the other is unrecognizable.
  You may assume that $\AllCFG$ is unrecognizable.
\end{problem}
\begin{Answer}
  Let's fix our alphabet to $\Sigma = \set{0, 1}$.
  \begin{enumerate}
    \item $\EqCFG$ is unrecognizable.
      We know that $\AllCFG$ is unrecognizable.
      Consider the following reduction of $\AllCFG$ to $\EqCFG$:

      \step
      $T$ = ``On input $\vector{M}$, where $M$ is a CFG:
        \begin{enumarabic}
          \item Construct PDA $P$ as follows:

            $P$ = ``On input $\vector{w}$:
              \begin{enumarabic}
                \item If $w \in \Sigma^*$, \Accept.
                \item Otherwise, \Reject.''
              \end{enumarabic}
          
          \item Construct CFG $G$ equivalent to $P$,
            such that $\cL(G) = \cL(P) = \Sigma^*$.
          \item \Return $\vector{M, G}$. ''
        \end{enumarabic}
      \begin{align*}
        M \in \AllCFG &\Iff M \text{ accepts all strings in the input alphabet.} \\
                      &\Iff \cL(M) = \Sigma^* \\ 
                      &\Iff \vector{M, G} \in \EqCFG
      \end{align*}
      Since $\AllCFG$ is unrecognizable, $\EqCFG$ must also be unrecognizable.
    \newpage
    \item $\EqCFGcomp$ is recognizable.
    
      \step
      Let $G_1$ and $G_2$ be two CFGs defined over the same alphabet
      $\Sigma$ and $M_1$ and $M_2$ be the equivalent PDAs
      such that $\vector{G_1, G_2} \in \EqCFGcomp$.
      This means $\cL(M_1)\neq \cL(M_2)$, so there exists \emph{at least}
      one string in $x \in \Sigma^*$ such that $x \in \cL(M_1)$
      or $x \in \cL(M_2)$, but not both
      (equivalently, $x \in \cL(M_1) \cup \cL(M_2)$
      and $x \notin \cL(M_1) \cap \cL(M_2)$).

      \step
      Since $\Sigma^*$ is decidable, we can construct a TM $E$ to
      list members of $\Sigma^*$. in lexicographic order.
      
      We can then recognize $\EqCFGcomp$ by the following TM $T$:

      \step
      $T$ = ``On input $\vector{M_1, M_2}$:
        \begin{enumarabic}
          \item Run $E$ to enumerate members of $\Sigma^*$.
          \item As each string is listed:
            \begin{enumarabic*}
              \item Simulate $M_1$ on $x$ and $M_2$ on $x$.
              \item If $M_1$ accepts $x$ and $M_2$ rejects $x$
                or $M_1$ rejects $x$ and $M_2$ accepts $x$,
                \Accept.
              \item Otherwise, continue enumerating strings and testing them.''
            \end{enumarabic*}
        \end{enumarabic}
  \end{enumerate}

  \step
  If an input $\vector{M_1, M_2}$ is in $\EqCFGcomp$,
  then $T$ will eventually halt and accept the input
  when it encounters a string not accepted by both $M_1$ and $M_2$
   r rejected by both $M_1$ and $M_2$.

  \step
  However, if an input $\vector{M_1, M_2}$ is not in $\EqCFGcomp$,
  then $T$ will never halt since $\Sigma^*$ is infinite.
\end{Answer}
