\begin{problem}
  Do the same for each of the following languages:
  \begin{enumalph}
    \item $L = \set{b_i \# b_{i+1} : i \geq 1}$, where $b$ is the binary
      representation of $i$ with no leading $0$'s.
      \begin{Answer}
        The language is not context-free. Suppose it was and $p$ as the pumping length.
        Take $s = 10^p1^p\#10^{p-1}10^p$. Since $10^p1^p + 1 = 10^{p-1}10^p$
        for all $p$, the string is guaranteed to be in the language.

        Write $s = uvwxy$ where, per the pumping lemma (\ref{lemma:pl}),
        $\abs{vwx} \leq p$, $\abs{vx} > 0$, and $uv^kwx^ky \in L$ for all $k \geq 0$.

        There are three possibilities:
        \begin{enumroman}
          \item If $vwx$ is a substring to the left of the $\#$ symbol in $s$,
            then, since $\abs{vx} > 0$, pumping down the string changes the left
            part of the string (previously equivalent to $b_i$)
            by removing \emph{at least} one symbol. Therefore, the resulting
            string cannot be equivalent to $b_i$. Since the right side of the string
            ($b_{i+1}$) remains unaltered, the new string is of the form
            $s' = b_j\#b_{i+1}$ for some $j \neq i$, so it is not in $L$,
            contradicting the pumping lemma.
          \item Similarly, if $vwx$ is a substring to the right of the $\#$ symbol,
            then pumping down the string changes the right part of the string
            (previously equivalent to $b_{i+1}$)
            by removing \emph{at least} one symbol. Therefore, the resulting
            string cannot be equivalent to $b_{i+1}$. Since the left side of the string
            ($b_i$) remains unaltered, the new string is of the form
            $s' = b_i\#b_j$ for some $j \neq i+1$, so it is not in $L$,
            contradicting the pumping lemma.
          \item If $vwx$ is a substring in the middle of the string, then:
            \begin{itemize}
              \item If $\#$ is contained in $\abs{vx}$, pumping down
                results in a string without a $\#$ symbol, so it is not in $L$.
              \item If $\#$ is contained in $\abs{w}$, then:
                \begin{itemize}
                  \item If the $1$ immediately after the $\#$ is not in $w$,
                    then it must be in $x$ so pumping down removes it
                    (but does not remove all the other $0$'s) resulting in a string
                    to the left that has leading $0$'s, hence cannot be equivalent
                    to $b_i$.
                  \newpage
                  \item If the $1$ immediately after the $\#$ in the original string
                    is in $w$, then $vx$ reads $1^a0^b$ for some $0 < a + b < p-2$.
                    Pumping down the string suggests that
                    $10^p1^{p-a}\#10^{p-1-b}10^p$ is in $L$, so
                    \begin{align*}
                      10^p1^{p-a} + 1 &= 10^{p-1-b}10^p \\
                      \Therefore 10^{\zaff{p-1}}10^{\crim{p - a}} &= 10^{\zaff{p-1-b}}10^{\crim{p}}\\
                      \Therefore p - a &= p &\Longrightarrow \quad a = 0 \\
                      \Therefore p - 1 - b &= p - 1 &\Longrightarrow  \quad b = 0
                    \end{align*}
                    But if $\abs{v} = a = 0$ and $\abs{x} = b = 0$, then
                    $\abs{vx} = 0$, which contradicts the pumping lemma.
                \end{itemize}
            \end{itemize}

            \step
            Therefore, $L$ must not be context-free since the pumping lemma is not 
            satisfied for some strings in $L$ of length greater than $p$.
        \end{enumroman}

      \end{Answer}
    \item $L = (a \cup b)^* - \set{(a^n b^n)^n : n \geq 1}$
    \begin{Answer}
      I got stuck on this problem.
    \end{Answer}
  \end{enumalph}
\end{problem}
